---
title: "SERI-LUC notebook"
output: html_document
date: "2024-12-10"
---

Load Packages
```{r Load packages, include=FALSE}
library(sf)
library(tidyverse)
library(ggplot2)
library(nomisr)
library(httr)
library(readxl)
library(biscale)
library(cowplot)
```

Load Data
```{r Load data, include=FALSE}
# Load 2021 MSOA boundary file and filter for England only
MSOA_21_boundary <- st_read("/dbfs/mnt/base/unrestricted/source_ons_open_geography_portal/dataset_middle_s_o_a_2021_ew_bfe/format_SHP_middle_s_o_a_2021_ew_bfe/LATEST_middle_s_o_a_2021_ew_bfe/MSOA_2021_EW_BFC_V7.shp") %>%
  filter(substr(MSOA21CD, 1, 1) == "E")

# Load 2011 MSOA boundary file and filter for England only
MSOA_11_boundary <- st_read("/dbfs/mnt/lab/unrestricted/elmantas.kamaitis@defra.gov.uk/Middle_layer_Super_Output_Areas_Dec_2011_Boundaries_Full_Clipped_BFC_EW_V3_2022_942885142986844588.gpkg") %>%
  filter(substr(MSOA11CD, 1, 1) == "E")

# Load Business Register and Employment Survey data on the number of workers per 4-digit 2007 SIC industry working in each 2011 MSOA
bres_sec_emp <- nomis_get_data(
  id = "NM_189_1",
  geography = "TYPE297",
  date = "latest",
  industry = "TYPE33",
  employment_status = 1,
  measure = 1,
  measures = 20100
)

# Load Census 2021 data on the number of employees per 3-digit SOC 2020 industry in each MSOA
census_sec_emp <- nomis_get_data(
  id = "NM_2081_1",
  geography = "TYPE152",
  c2021_occ_105 = "0...104",
  measures = 20100
)

# Load ONS experimental occupational skills dataset
temp_file <- tempfile(fileext = ".xlsx")

download.file("https://www.ons.gov.uk/file?uri=/employmentandlabourmarket/peopleinwork/labourproductivity/datasets/skillssupplyestimates2012to2023/august2024/skillssupply2012to2023.xlsx", temp_file, mode = "wb") # onsr does not have this dataset available, so loading it from link instead

ons_occ_skills <- read_excel(temp_file, sheet = 4, skip = 4)
```
##### BRES based analysis #####

Summing by industry and merging with MSOA boundaries
```{r}
# Summing OBS_VALUE for each unique GEOGRAPHY
summed_df <- bres_sec_emp %>%
  group_by(GEOGRAPHY, GEOGRAPHY_CODE) %>%
  summarise(SUM_OBS_TOTAL = sum(as.numeric(OBS_VALUE), na.rm = TRUE), .groups = 'drop')

# Summing OBS_VALUE for all 4-digit INDUSTRY_CODE that start with "01" and are therefore in agri-industries for each unique GEOGRAPHY
summed_industry_df <- bres_sec_emp %>%
  filter(substr(INDUSTRY_CODE, 1, 2) == "01") %>%
  group_by(GEOGRAPHY) %>%
  summarise(SUM_OBS_AGRI = sum(as.numeric(OBS_VALUE), na.rm = TRUE))

# Merge the two dataframes to combine the results
final_df <- left_join(summed_df, summed_industry_df, by = "GEOGRAPHY")

# Calculate the percentage of SUM_OBS_AGRI in SUM_OBS_TOTAL for each GEOGRAPHY
final_df <- final_df %>%
  mutate(AGRI_PERCENTAGE = (SUM_OBS_AGRI / SUM_OBS_TOTAL) * 100) %>%
  filter(substr(GEOGRAPHY_CODE, 1, 1) == "E")

# Merge the employment sums data with the MSOA boundary file
merged_data <- left_join(MSOA_11_boundary, final_df, by = c("MSOA11CD" = "GEOGRAPHY_CODE"))

```
Summary Stats
```{r}

summary(merged_data$AGRI_PERCENTAGE)

summary(merged_data$SUM_OBS_TOTAL)

total_sum_obs <- merged_data %>%
  st_drop_geometry() %>%  # Drop geometry to simplify the operation
  summarise(Total_SUM_OBS_TOTAL = sum(SUM_OBS_TOTAL, na.rm = TRUE))

# Create a histogram for AGRI_PERCENTAGE
ggplot(data = merged_data, aes(x = AGRI_PERCENTAGE)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(
    title = "Histogram of Agriculture Percentage",
    x = "Agriculture Percentage",
    y = "Frequency"
  ) +
  theme_minimal()


# Process and bin AGRI_PERCENTAGE, dropping geometry and unnecessary columns
bin_counts <- merged_data %>%
  st_drop_geometry() %>%  # Remove geometry column
  mutate(
    AGRI_PERCENTAGE_BIN = case_when(
      AGRI_PERCENTAGE == 0 ~ "0",  # Bin for exactly 0 values
      AGRI_PERCENTAGE > 0 & AGRI_PERCENTAGE <= 0.5 ~ "Near Zero (0 to 0.5)",  # Bin for near-zero values
      TRUE ~ cut(AGRI_PERCENTAGE, breaks = seq(0.5, 100, by = 0.5), include.lowest = TRUE, right = FALSE)  # Other bins
    )
  ) %>%
  count(AGRI_PERCENTAGE_BIN, name = "Count") %>%  # Count occurrences in each bin
  arrange(AGRI_PERCENTAGE_BIN)  # Sort bins

# Print the resulting table
print(bin_counts)

```
Map plots for percentage and total agri employment
```{r}

##### Total employment plot #####
# Calculate the threshold for the upper 5% of SUM_OBS_TOTAL
threshold <- quantile(merged_data$SUM_OBS_TOTAL, 0.95, na.rm = TRUE)

# Add a new column to classify regions as top 5% or not
merged_data <- merged_data %>%
  mutate(SUM_OBS_CATEGORY = ifelse(SUM_OBS_TOTAL >= threshold, "Top 5%", "Other"))

# Plot the data 5% excluded total employment data
ggplot(data = merged_data) +  
  geom_sf(data = merged_data %>% filter(SUM_OBS_CATEGORY == "Other"), # Separate the fill aesthetic for the two categories
          aes(fill = ifelse(SUM_OBS_TOTAL == 0, NA, SUM_OBS_TOTAL)), color = "grey", size = 0.1) +
  geom_sf(data = merged_data %>% filter(SUM_OBS_CATEGORY == "Top 5%"), 
          fill = "red", color = "grey", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", na.value = "white", name = "Employment Count") +  # Continuous fill scale for values excluding top 5%
  theme_minimal() +
  labs(title = "Total Employment by MSOA",
       subtitle = "Top 5% areas highlighted in red") +
  theme(legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 48, hjust = 0.5),      # Adjust title size
        plot.subtitle = element_text(size = 44, hjust = 0.5),   # Adjust subtitle size
        legend.title = element_text(size = 32),                 # Adjust legend title size
        legend.text = element_text(size = 30),                  # Adjust legend text size
        axis.text = element_text(size = 30),                    # Adjust axis text size
        legend.key.height = unit(4.5, "cm"),                    # Adjust the height of the legend key
        legend.key.width = unit(4.5, "cm")                      # Adjust the width of the legend key
  )

# Save the plot
ggsave("employ_5pct_total_MSOA.png", device = "png", width = 45, height = 45, dpi = 200)

##### Agriculture proportion employment #####
# Map agriculture proportion of total employment
ggplot(data = merged_data) +
  geom_sf(aes(fill = ifelse(AGRI_PERCENTAGE == 0, NA, AGRI_PERCENTAGE)), color = "grey", size = 0.001) + 
  # Boundary lines are grey, areas with 0 are set to NA
  scale_fill_viridis_c(option = "plasma", na.value = "white", name = "Agriculture %") +  
  # White for areas with 0 values
  theme_minimal() +
  labs(title = "Agricultural related employment percentage by 2011 MSOA",
       subtitle = "Agricultural (SIC 2007 01**) employment as a percentage of total employment using BRES data") +
  theme(legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 48, hjust = 0.5),      # Adjust title size
        plot.subtitle = element_text(size = 44, hjust = 0.5),   # Adjust subtitle size
        legend.title = element_text(size = 32),                 # Adjust legend title size
        legend.text = element_text(size = 30),                  # Adjust legend text size
        axis.text = element_text(size = 30),                    # Adjust axis text size
        legend.key.height = unit(4.5, "cm"),                    # Adjust the height of the legend key
        legend.key.width = unit(4.5, "cm")                      # Adjust the width of the legend key
        )

# Save the plot
ggsave("agri_percentage_BRES.png", device = "png", width = 45, height = 45, dpi = 200)

##### Agriculture total employment #####
# Map agriculture employment
ggplot(data = merged_data) +
  geom_sf(aes(fill = ifelse(SUM_OBS_AGRI == 0, NA, SUM_OBS_AGRI)), color = "grey", size = 0.001) + 
  # Boundary lines are grey, areas with 0 are set to NA
  scale_fill_viridis_c(option = "plasma", na.value = "white", name = "Agriculture Employment Count") +  
  # White for areas with 0 values
  theme_minimal() +
  labs(title = "Agricultural Employment by MSOA",
       subtitle = "Agricultural employment by MSOA in England") +
  theme(legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

# Save the plot
ggsave("agri_total_MSOA.png", device = "png", width = 45, height = 45, dpi = 200)

# Calculate the threshold for the upper 0.1% of SUM_OBS_TOTAL
threshold <- quantile(merged_data$SUM_OBS_AGRI, 0.999, na.rm = TRUE)

# Add a new column to classify regions as top 0.1% or not
merged_data <- merged_data %>%
  mutate(SUM_OBS_CATEGORY = ifelse(SUM_OBS_AGRI >= threshold, "Top 0.1%", "Other"))

# Plot the data 0.5% excluded total employment data
ggplot(data = merged_data) +  
  geom_sf(data = merged_data %>% filter(SUM_OBS_CATEGORY == "Other"), # Separate the fill aesthetic for the two categories
          aes(fill = ifelse(SUM_OBS_AGRI == 0, NA, SUM_OBS_AGRI)), color = "grey", size = 0.1) +
  geom_sf(data = merged_data %>% filter(SUM_OBS_CATEGORY == "Top 0.1%"), 
          fill = "red", color = "grey", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", na.value = "white", name = "Agri employment Count") +  # Continuous fill scale for values excluding top 5%
  theme_minimal() +
  labs(title = "Agri employment by MSOA",
       subtitle = "Top 0.1% areas highlighted in red") +
  theme(legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 48, hjust = 0.5),      # Adjust title size
        plot.subtitle = element_text(size = 44, hjust = 0.5),   # Adjust subtitle size
        legend.title = element_text(size = 32),                 # Adjust legend title size
        legend.text = element_text(size = 30),                  # Adjust legend text size
        axis.text = element_text(size = 30),                    # Adjust axis text size
        legend.key.height = unit(4.5, "cm"),                    # Adjust the height of the legend key
        legend.key.width = unit(4.5, "cm")                      # Adjust the width of the legend key
  )

# Save the plot
ggsave("agri_01pct_total_MSOA.png", device = "png", width = 45, height = 45, dpi = 200)


```

##### Census based analysis #####

Plot proportion Agri (511) of total employment
```{r}
# Filter and summarize census data for the required categories
filtered_census_data <- census_sec_emp %>%
  filter(C2021_OCC_105 %in% c(0, 60)) %>%   # Retain rows where C2021_OCC_105 is 0 or 60
  mutate(Category = case_when(
    C2021_OCC_105 == 0 ~ "SUM_OBS_TOTAL",  # Rename categories
    C2021_OCC_105 == 60 ~ "SUM_OBS_AGRI"
  )) %>%
  group_by(GEOGRAPHY_CODE, Category) %>%    # Group by geography and category
  summarise(SUM_OBS_VALUE = sum(OBS_VALUE, na.rm = TRUE), .groups = "drop") %>% 
  pivot_wider(names_from = Category, values_from = SUM_OBS_VALUE, values_fill = 0)  # Reshape to wide format

# Merge the summarised census data with the shapefile
census_merged_data <- MSOA_21_boundary %>%
  left_join(filtered_census_data, by = c("MSOA21CD" = "GEOGRAPHY_CODE")) %>%
  mutate(
    AGRI_PERCENTAGE = ifelse(SUM_OBS_TOTAL > 0, SUM_OBS_AGRI / SUM_OBS_TOTAL * 100, NA)  # Calculate percentage
  )

# Map agriculture proportion of total employment
ggplot(data = census_merged_data) +
  geom_sf(aes(fill = ifelse(AGRI_PERCENTAGE == 0, NA, AGRI_PERCENTAGE)), color = "grey", size = 0.001) + 
  # Boundary lines are grey, areas with 0 are set to NA
  scale_fill_viridis_c(option = "plasma", na.value = "white", name = "Agriculture %") +  
  # White for areas with 0 values
  theme_minimal() +
  labs(title = "Agricultural and Related Trades by 2021 MSOA",
       subtitle = "Agricultural and Related Trades as a percentage of total employment using Census 2021 data") +
  theme(legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 48, hjust = 0.5),      # Adjust title size
        plot.subtitle = element_text(size = 44, hjust = 0.5),   # Adjust subtitle size
        legend.title = element_text(size = 32),                 # Adjust legend title size
        legend.text = element_text(size = 30),                  # Adjust legend text size
        axis.text = element_text(size = 30),                    # Adjust axis text size
        legend.key.height = unit(4.5, "cm"),                    # Adjust the height of the legend key
        legend.key.width = unit(4.5, "cm")                      # Adjust the width of the legend key
  )

# Save the plot
ggsave("agri_percentage_census.png", device = "png", width = 45, height = 45, dpi = 200)

```

Estimate the leading employment industries in each MSOA
```{r}

# Estimate the proportion of total employment for each industry in the census data
processed_data <- census_sec_emp %>%
  filter(str_starts(GEOGRAPHY_CODE, "E")) %>%  # Filter for GEOGRAPHY_CODE starting with "E"
  group_by(GEOGRAPHY_CODE) %>%  # Group by each unique GEOGRAPHY_CODE
  mutate(
    # Get the OBS_VALUE where C2021_OCC_105 is 0 for each GEOGRAPHY_CODE
    Total_OBS_VALUE = OBS_VALUE[C2021_OCC_105 == 0],
    # Divide each OBS_VALUE by the Total_OBS_VALUE
    PERC_OBS_VALUE = ifelse(!is.na(Total_OBS_VALUE) & Total_OBS_VALUE > 0, OBS_VALUE / Total_OBS_VALUE, NA)
  ) %>%
  ungroup()  # Ungroup after processing

# Extract the top 3 values for each GEOGRAPHY_CODE, excluding rows where C2021_OCC_105 == 0
top_3_values <- processed_data %>%
  group_by(GEOGRAPHY_CODE) %>%
  filter(!is.na(PERC_OBS_VALUE) & C2021_OCC_105 != 0) %>%  # Exclude the overall total
  arrange(desc(PERC_OBS_VALUE)) %>%
  slice_max(order_by = PERC_OBS_VALUE, n = 3, with_ties = FALSE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Reshape the data to wide format, creating separate columns for the top 3 values and their names
top_3_wide <- top_3_values %>%
  select(GEOGRAPHY_CODE, Rank, PERC_OBS_VALUE, C2021_OCC_105_NAME) %>%
  pivot_wider(
    names_from = Rank,
    values_from = c(PERC_OBS_VALUE, C2021_OCC_105_NAME),
    names_prefix = "Top_"
  )

# Merge the result with MSOA_21_boundary
top_prop_employ <- MSOA_21_boundary %>%
  inner_join(top_3_wide, by = c("MSOA21CD" = "GEOGRAPHY_CODE"))

summary(top_prop_employ$PERC_OBS_VALUE_Top_1)

# Count appearances of unique values in C2021_OCC_105_NAME_Top_1
top_1_employ_counts <- top_prop_employ %>%
  st_drop_geometry() %>%  # Drop the geometry column
  count(C2021_OCC_105_NAME_Top_1, name = "Count") %>%
  arrange(desc(Count))

# View the sorted table
print(top_1_employ_counts)

# Count appearances of unique values in C2021_OCC_105_NAME_Top_2
top_2_employ_counts <- top_prop_employ %>%
  st_drop_geometry() %>%  # Drop the geometry column
  count(C2021_OCC_105_NAME_Top_2, name = "Count") %>%
  arrange(desc(Count))

# View the sorted table
print(top_2_employ_counts)

# Count appearances of unique values in C2021_OCC_105_NAME_Top_3
top_3_employ_counts <- top_prop_employ %>%
  st_drop_geometry() %>%  # Drop the geometry column
  count(C2021_OCC_105_NAME_Top_3, name = "Count") %>%
  arrange(desc(Count))

# View the sorted table
print(top_3_employ_counts)

```

Plot the MSOAs where agri (511) is the leading employment industry
```{r}
# Merge top_3_values with MSOA_21_boundary to include geometry
top_3_with_geometry <- top_3_values %>%
  inner_join(MSOA_21_boundary, by = c("GEOGRAPHY_CODE" = "MSOA21CD"))

# Filter for C2021_OCC_105_CODE == "_60"
filtered_top_3 <- top_3_with_geometry %>%
  filter(C2021_OCC_105_CODE == "_60")  # Keep rows where code is _60

# Plot the filtered data
ggplot(data = filtered_top_3) +
  geom_sf(aes(geometry = geometry, color = factor(Rank)), size = 0.5) +  # Use different colors for Rank
  scale_color_manual(
    values = c("red", "blue", "green"),  # Assign colors to Rank 1, 2, and 3
    name = "Rank",
    labels = c("Rank 1", "Rank 2", "Rank 3")  # Legend labels
  ) +
  theme_minimal() +
  labs(
    title = "Top 3 Ranks for C2021_OCC_105_CODE == _60",
    subtitle = "Ranks displayed in different colors",
    color = "Rank"
  ) +
  theme(
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )


# Plot the filtered data with boundaries underneath
ggplot() +
  # Add grey boundaries from MSOA_21_boundary
  geom_sf(data = MSOA_21_boundary, fill = "transparent", colour = "grey", size = 0.5) + 
  # Add the filtered top 3 ranked locations
  geom_sf(data = filtered_top_3, aes(geometry = geometry, fill = factor(Rank)), size = 0.5) + 
  scale_fill_manual(
    values = c("red4", "orange4", "yellow4"),  # Assign colors to Rank 1, 2, and 3
    name = "Rank",
    labels = c("Rank 1", "Rank 2", "Rank 3")  # Legend labels
  ) +
  theme_minimal() +
  labs(
    title = "Agricultural and Related Trades main proportional employment industry",
    subtitle = "2021 MSOAs where Agricultural and Related Trades is the top 3 proportional employment industry using Census 2021 data",
    color = "Rank"
  ) +
  theme(legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 48, hjust = 0.5),      # Adjust title size
        plot.subtitle = element_text(size = 44, hjust = 0.5),   # Adjust subtitle size
        legend.title = element_text(size = 32),                 # Adjust legend title size
        legend.text = element_text(size = 30),                  # Adjust legend text size
        axis.text = element_text(size = 30),                    # Adjust axis text size
        legend.key.height = unit(4.5, "cm"),                    # Adjust the height of the legend key
        legend.key.width = unit(4.5, "cm")                      # Adjust the width of the legend key
  )

ggsave("top3_employ_census.png", device = "png", width = 45, height = 45, dpi = 200)
```

Skill matching for Agri (511) industry
```{r}
# Function to compare skills for all occupations against a given SOC 2020 code
compare_all_skills <- function(data, reference_soc_code) {
  # Filter skills for the reference SOC 2020 code
  reference_skills <- data %>%
    filter(`SOC 2020 code` == reference_soc_code) %>%
    select(`Skill O*NET identifier`, `Average competence level (0 to 7 scale)`) %>%
    rename(`reference_level` = `Average competence level (0 to 7 scale)`)
  
  # Join the reference skills with the full dataset
  comparison <- data %>%
    inner_join(reference_skills, by = "Skill O*NET identifier") %>%
    mutate(is_equal_or_lower = `Average competence level (0 to 7 scale)` <= reference_level,
           is_within_5_percent = `Average competence level (0 to 7 scale)` <= reference_level * 1.05,
           is_within_10_percent = `Average competence level (0 to 7 scale)` <= reference_level * 1.10) %>%
    distinct(`SOC 2020 code`, `SOC 2020 unit group title`, `Skill O*NET identifier`, .keep_all = TRUE)  # Ensure unique skill per occupation
  
  # Summarize the results for each occupation
  result <- comparison %>%
    group_by(`SOC 2020 code`, `SOC 2020 unit group title`) %>%
    summarize(
      equal_or_lower_skill_count = sum(is_equal_or_lower),
      within_5_percent_skill_count = sum(is_within_5_percent),
      within_10_percent_skill_count = sum(is_within_10_percent),
      .groups = "drop"
    )
  
  # Rank the occupations based on the number of matching skills
  result <- result %>%
    mutate(
      rank_equal_or_lower = rank(-equal_or_lower_skill_count, ties.method = "min"),
      rank_within_5_percent = rank(-within_5_percent_skill_count, ties.method = "min"),
      rank_within_10_percent = rank(-within_10_percent_skill_count, ties.method = "min"),
      # Create the 3-digit SOC code by removing the last digit of the SOC 2020 code
      `3 digit SOC 2020` = substr(as.character(`SOC 2020 code`), 1, 3)
    )
  
  # Calculate the average skill counts for each 3-digit SOC code
  result <- result %>%
    group_by(`3 digit SOC 2020`) %>%
    mutate(
      avg_equal_or_lower_skill_count = mean(equal_or_lower_skill_count, na.rm = TRUE),
      avg_within_5_percent_skill_count = mean(within_5_percent_skill_count, na.rm = TRUE),
      avg_within_10_percent_skill_count = mean(within_10_percent_skill_count, na.rm = TRUE)
    ) %>%
    ungroup()  # Remove grouping
  
  return(result)
}

# Run the function with SOC 2020 code 5111 for the "Farmers" occupation
farmers_skill_level <- compare_all_skills(ons_occ_skills, 5111)

```

Identify the top 10 highest average skill compatible occupation (3-digit) groups
```{r}
# Sort the result by avg_equal_or_lower_skill_count in descending order and select the top 10, excluding other agricultural SOCs
top_10_soc_group <- compare_all_skills(ons_occ_skills, 5111) %>%
  arrange(desc(avg_equal_or_lower_skill_count)) %>%
  filter(!`3 digit SOC 2020` %in% c("511", "911")) %>% # Filter out other Agri codes
  distinct(`3 digit SOC 2020`, .keep_all = TRUE) %>%
  slice_head(n = 10) %>%
  select(`3 digit SOC 2020`, avg_equal_or_lower_skill_count)

# Extract the first 3 characters of C2021_OCC_105_NAME to match with `3 digit SOC 2020`
census_sec_emp_selected <- census_sec_emp %>%
  mutate(`3 digit SOC 2020` = substr(C2021_OCC_105_NAME, 1, 3)) %>%
  select(`3 digit SOC 2020`, C2021_OCC_105) %>%
  distinct(`3 digit SOC 2020`, .keep_all = TRUE) # only keep unique values

# Merge the `top_10` list with the selected `C2021_OCC_105` column
top_10_occ_105_group <- top_10_soc_group %>%
  left_join(census_sec_emp_selected, by = "3 digit SOC 2020")

# Print the matched results
print(top_10_occ_105_group)
```

Plot the total employment in the top 10 skill-matching occ groups
```{r}
# Convert both columns to character and filter matching occupations
matched_census_data <- census_sec_emp %>%
  mutate(C2021_OCC_105 = as.character(C2021_OCC_105)) %>%
  semi_join(
    top_10_occ_105_group %>% mutate(C2021_OCC_105 = as.character(C2021_OCC_105)), 
    by = "C2021_OCC_105"
  )

# Summarize total OBS_VALUE for matching occupations and overall employment (C2021_OCC_105 = 0)
top_occ_sum_per_geography <- census_sec_emp %>%
  mutate(C2021_OCC_105 = as.character(C2021_OCC_105)) %>%
  group_by(GEOGRAPHY_CODE) %>%
  summarize(
    top_occ_obs_value = sum(OBS_VALUE[C2021_OCC_105 %in% top_10_occ_105_group$C2021_OCC_105], na.rm = TRUE),
    total_obs_value = sum(OBS_VALUE[C2021_OCC_105 == "0"], na.rm = TRUE),
    agri_obs_value = sum(OBS_VALUE[C2021_OCC_105 == "60"], na.rm = TRUE),
    top_percentage_value = ifelse(
      total_obs_value > 0, 
      (top_occ_obs_value / total_obs_value) * 100, 
      NA_real_  # Assign NA if total_obs_value is 0
    ),
    agri_percentage_value = ifelse(
      agri_obs_value > 0, 
      (agri_obs_value / total_obs_value) * 100, 
      0  # Assign NA if total_obs_value is 0
    ),
    agri_top_ratio = ifelse(
      agri_obs_value > 0, 
      (agri_obs_value / top_occ_obs_value), 
      NA_real_  # Assign NA if agri_obs_value is 0
    ),
    .groups = "drop"
  ) %>%
  filter(substr(GEOGRAPHY_CODE, 1, 1) == "E")

# Merge top_occ_sum_per_geography with MSOA_21_boundary based on GEOGRAPHY_CODE and MSOA21CD
top_occ_sum_geo_sf <- top_occ_sum_per_geography %>%
  left_join(MSOA_21_boundary, by = c("GEOGRAPHY_CODE" = "MSOA21CD")) %>%
  st_as_sf()

# Plot the top 10 skill-matching industry total employment by MSOA
ggplot(data = top_occ_sum_geo_sf) +  
  geom_sf(aes(geometry = geometry, fill = ifelse(top_occ_obs_value == 0, NA, top_occ_obs_value)), color = "grey", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", na.value = "white", name = "Employment Count") +  # Continuous fill scale for values
  theme_minimal() +
  labs(title = "Total Employment in top 10 skill matching industries by MSOA",
       subtitle = "Skills matched based on ONS' 'Important skills required per occupation'") +
  theme(legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 48, hjust = 0.5),      # Adjust title size
        plot.subtitle = element_text(size = 44, hjust = 0.5),   # Adjust subtitle size
        legend.title = element_text(size = 32),                 # Adjust legend title size
        legend.text = element_text(size = 30),                  # Adjust legend text size
        axis.text = element_text(size = 30),                    # Adjust axis text size
        legend.key.height = unit(4.5, "cm"),                    # Adjust the height of the legend key
        legend.key.width = unit(4.5, "cm")                      # Adjust the width of the legend key
  )

# Save the plot
ggsave("skill_match_plot.png", device = "png", width = 45, height = 45, dpi = 200)

# Plot the top 10 skill-matching industry as proportion of total employment by MSOA
ggplot(data = top_occ_sum_geo_sf) +  
  geom_sf(aes(geometry = geometry, fill = ifelse(top_percentage_value == 0, NA, top_percentage_value)), color = "grey", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", na.value = "white", name = "Proportion of total employment") +  # Continuous fill scale for values
  theme_minimal() +
  labs(title = "Proportional employment in top 10 skill matching industries by MSOA",
       subtitle = "Skills matched based on ONS' 'Important skills required per occupation'") +
  theme(legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 48, hjust = 0.5),      # Adjust title size
        plot.subtitle = element_text(size = 44, hjust = 0.5),   # Adjust subtitle size
        legend.title = element_text(size = 32),                 # Adjust legend title size
        legend.text = element_text(size = 30),                  # Adjust legend text size
        axis.text = element_text(size = 30),                    # Adjust axis text size
        legend.key.height = unit(4.5, "cm"),                    # Adjust the height of the legend key
        legend.key.width = unit(4.5, "cm")                      # Adjust the width of the legend key
  )

# Save the plot
ggsave("skill_match_percentage.png", device = "png", width = 45, height = 45, dpi = 200)

# Plot the top 10 skill-matching industry as proportion of total employment by MSOA
ggplot(data = top_occ_sum_geo_sf) +  
  geom_sf(aes(geometry = geometry, fill = ifelse(agri_top_ratio == 0, NA, agri_top_ratio)), color = "grey", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", na.value = "white", name = "Agri/Top 10 skill-matched industry employment ratio") +  # Continuous fill scale for values
  theme_minimal() +
  labs(title = "Ratio between Agri (511) and top 10 skill matching industries by MSOA",
       subtitle = "Skills matched based on ONS' 'Important skills required per occupation'") +
  theme(legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 48, hjust = 0.5),      # Adjust title size
        plot.subtitle = element_text(size = 44, hjust = 0.5),   # Adjust subtitle size
        legend.title = element_text(size = 32),                 # Adjust legend title size
        legend.text = element_text(size = 30),                  # Adjust legend text size
        axis.text = element_text(size = 30),                    # Adjust axis text size
        legend.key.height = unit(4.5, "cm"),                    # Adjust the height of the legend key
        legend.key.width = unit(4.5, "cm")                      # Adjust the width of the legend key
  )

# Save the plot
ggsave("skill_match_ratio.png", device = "png", width = 45, height = 45, dpi = 200)

```

Agri (511) and top 10 skill-matched occ group Bivariate plot
```{r}
# Classify data for the bivariate map
top_occ_sum_geo_sf <- top_occ_sum_geo_sf %>%
  bi_class(
    x = agri_percentage_value,
    y = top_percentage_value,
    style = "quantile",  # or "equal"
    dim = 3              # 3x3 grid for 9 bivariate color categories
  )

# Plot the bivariate choropleth map
bivariate_map <- ggplot(top_occ_sum_geo_sf) +
  geom_sf(aes(geometry = geometry, fill = bi_class), color = "grey", size = 0.1) +
  bi_scale_fill(pal = "DkViolet", dim = 3) +  # Choose a color palette
  labs(
    title = "Bivariate plot of Agriculture (SOC 511) vs Top 10 Skill-Matched Industry Employment (%)",
    subtitle = "Skill matched industries exclude similar agricultural industries (SOC 511, 911)",
    caption = "Source: Census Employment Data, ONS experimental occupational skills data"
  ) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 48, hjust = 0.5),
    plot.subtitle = element_text(size = 44, hjust = 0.5),
    plot.caption = element_text(size = 24, hjust = 0.5),
  ) + 
    guides(fill = "none")

# Add a legend to the map
bivariate_legend <-
  bi_legend(
    pal = "DkViolet",
    dim = 3,
    xlab = "Higher Agricultural %",
    ylab = "Higher S-M Industry %",
    size = 32  # Adjust legend size
  ) +
  theme(
    plot.background = element_rect(fill = "grey90"),
    panel.background = element_rect(fill = "grey90"))

# Combine the map and legend using cowplot
final_plot <- cowplot::plot_grid(
  bivariate_map, 
  bivariate_legend, 
  ncol = 2,         # Place map and legend side by side
  rel_widths = c(4, 1)  # Adjust the relative widths of the map and legend
)

# Save the plot
ggsave("bi_chorop_agri_skill_match_v2.png", device = "png", width = 45, height = 45, dpi = 200)

```
Generate an interactive bivariate Agri (511) and top 10 skill-matched occ group leaflet plot  
```{r}

# Ensure geometry is valid and CRS is correct
top_occ_sum_geo_sf <- top_occ_sum_geo_sf %>%
  st_make_valid() %>%
  st_transform(crs = 4326)

# Simplify geometry for rendering
top_occ_sum_geo_sf <- st_simplify(top_occ_sum_geo_sf, dTolerance = 50)

# Define the colour palette
bivariate_colors <- c(
  "1-1" = "#cabed0",
  "1-2" = "#89a1c8",
  "1-3" = "#4885c1",
  "2-1" = "#bc7c8f",
  "2-2" = "#806a8a",
  "2-3" = "#435786",
  "3-1" = "#ae3a4e",
  "3-2" = "#77324c",
  "3-3" = "#3f2949"
)

# Match colour palette with bivariate classes 
top_occ_sum_geo_sf <- top_occ_sum_geo_sf %>%
  mutate(
    color_values = ifelse(
      !is.na(bi_class) & bi_class %in% names(bivariate_colors),
      bivariate_colors[bi_class],
      "grey90"  # Fallback for missing or invalid classes
    )
  )

# Generate the leaflet plot
map <- leaflet(top_occ_sum_geo_sf) %>%
  addPolygons(
    fillColor = ~color_values,
    weight = 0.5,
    color = "grey90",
    fillOpacity = 0.7,
    popup = ~paste(
      "<b>MSOA:</b>", MSOA21NM, "<br>",
      "<b>Agriculture %:</b>", round(agri_percentage_value, 1), "<br>",
      "<b>Top 10 Industry %:</b>", round(top_percentage_value, 1)
    )
  )

# Save and open in browser
htmlwidgets::saveWidget(map, "interactive_map.html")
browseURL("interactive_map.html")
```

